\mytitle

### 2021-10-26 Jason.tex
Update README.
Edits to get it ready for the next submission up to Sect 3.5.1,

\begin{abstract} 
%Wind farms play an increasingly important
%role as a source of renewable energy and 
  A key aspect of the design of a wind farm is the wind farm layout optimization (WFLO) problem: given a wind farm site and information about the wind patterns, the problem is to decide the location of individual wind turbines to maximize energy production subject to proximity restrictions and wake-based interference between turbines. Given the pairwise wake interactions, it is natural
%While a number of heuristic and exact techniques have been proposed to solve WFLO, it is common
to model the energy objective as a quadratic function as, indeed, has been done in some existing optimization approaches. However, state-of-the-art solutions often trade-off between speed in producing designs and quality in terms of finding and proving optimal solutions. In this work, we aim to find a balancing approach to obtain WFLO solutions quickly for interactive design \emph{and} solve the problem to optimality when quality is more important. To this end, we exploit recent advances in optimization hardware and software that target quadratic constraints: commercial mixed integer linear solvers have been extended to address some quadratic problems and nascent specialized hardware, including quantum computing systems, have focused on solving quadratic unconstrained binary optimization (QUBO) problems. 
%It is a key problem 
%when designing a new wind farm, as it has a direct influence
%on the amount of the total produced energy. 
%In recent years there have been advances in both software technologies (e.g., the extension of commercial mixed integer linear programming solvers to handle quadratic problems) and nascent hardware development (e.g., quantum computing chips and specialized classical chips) that have relevance for the WFLO problem. In both cases, the advances have 
%Recent years saw 
%advances in software and hardware optimization technologies, such as Gurobi's quadratic solver 
%and Fujitsu's digital annealer (DA), respectively. 
%These technologies create an opportunity to improve 
%over current wind farm optimization solutions
%both in terms of solution quality and solve time. 
%In order to exploit these technologies, 
%we must represent the WFLO problem as 
%a quadratic optimization model. Specifically, in order 
%to apply a software-based optimization
%solution (Gurobi), 
We 
introduce two novel quadratic programming models for WFLO: a quadratic constrained optimization problem (QCOP) with binary decision variables and a QUBO.
%Subsequently, we provide the unconstrained counterpart of the QCOP, 
%namely the quadratic unconstrained binary optimization (QUBO), which  
%is the underlying representation for quantum-inspired 
%optimization hardware such as the DA. 
A thorough numerical 
evaluation using a commercial solver and specialized QUBO hardware
%demonstrates the strengths of the two models. 
%Our results
show that our quadratic framework achieves fast, high-quality solutions 
that
improve the state of the art and strike a balance between speed and quality. In particular, the QUBO model delivers high quality solutions in a few seconds while the QCOP model can be used to find better solutions and provide quality guarantees over a longer run-time.
\end{abstract}
\begin{keyword}
	Wind Farm Layout Optimization (WFLO) \sep Quadratic Programming (QP) \sep Quadratic Unconstrained Binary Optimization (QUBO) \sep Digital Annealer
\end{keyword}

@@

\section{Introduction}
% Todo
The goal in the wind farm layout optimization (WFLO)
problem~\cite{MOSETTI1994105} is to locate wind turbines within a
predefined area to maximize the total power harvested from the wind
stream.  While physical considerations based on the diameter of the
blades prevent turbines being placed too close to each other, turbine
location decisions influence the total power production due to
interference effects (i.e., wakes) generated by upstream turbines.
These inter-turbine effects are best captured by Jensen's wake model~\cite{jensen1983note} when superimposed
using a sum-of-squares expression~\cite{Zhang2014}. However,
this wake modeling yields intractable optimization problems
that must be approximated using either quadratic or linear functions~\cite{turner2014new}.
An alternative formulation of WFLO relies on a less accurate wake model, using linear superpositioning
to capture multiple wakes~\cite{donovan2005wind}.
The latter can be captured using quadratic
constraints and objective functions. However, exact optimization
approaches that have been proposed for the alternative WFLO formulation~\cite{Zhang2014,kuo2016wind,donovan2005wind,fagerfjall2010optimizing,archer2011wind,sorkhabi2018constrained}
represent the quadratic relations using approximate linear models,
%that approximate wake dependencies,
as solving quadratic programming problems
is computationally challenging compared to linear models.

@@

In this work, we introduce two novel quadratic WFLO models with the goal 
of combining efficiency, physical farm representation, and optimality. The first
model represents WFLO as a quadratic constrained optimization
problem (QCOP).  The QCOP can be solved using state-of-the-art mathematical software
that has been recently extended for such problems, e.g., Gurobi~\cite{gurobi}.  Our
second model is a quadratic unconstrained binary optimization
(QUBO) representation of the WFLO problem.  Such a formulation enables the use of
nascent specialized optimization hardware tailored to solve QUBOs: we use Fujitsu's Digital Annealer (DA) \cite{aramon2019physics}. By employing the two types of advances (software and hardware) for solving
quadratic problems, 
we are able to provide fast and high performing solutions (in terms of the total energy). Moreover, our approach is based
on a physical model of the wind farm, taking into account
exact calculations of the total energy and the wake effects. 
Therefore, our approach achieves the desired balance between efficiency and effectiveness of the resulting design, in a model-based fashion.
%, a hardware-based
%QUBO solver, to demonstrate the strength of our approach.

@@

Our experiments are focused on the performance of the two approaches
compared to 
existing optimization methods. To this end,
we experiment with 12 standard WFLO benchmark problems (c.f.~\cite{turner2014new})
and show that solving quadratic WFLO models
using Gurobi and the DA
achieves state-of-the-art performance.
%, often outperforming existing solutions when provided sufficient run-time.
Furthermore, we observe that the DA finds solutions of
similar quality to software-based approaches but with two orders of
magnitude shorter run-time.
%The ability to quickly obtain competitive
%solutions on large WFLO instances can be employed to solve multiple
%wind farm design problems.

@@

We start this section by presenting the physical model of a wind farm
%used throughout the paper.
%Our wind farm model is
based on the common
notation and description presented in Zhang et al.\ \cite{Zhang2014}.  Subsequently,
we provide a literature review of existing mathematical optimization
methods for solving WFLO, focusing on their similarities and
differences with our approach.

@@

%\paragraph{Wind Farm as a Grid}
We consider a square wind farm that we model as an $n\times n$
grid with $n$ being the number of cells on each axis.  The cells have
an area of $c^2$ (in square meters), i.e., if $c=100$ meters and
$n=20$, the farm area will be $2 \mbox{ km} \times 2 \mbox{ km}$.  We
assume that the terrain is flat and that we have no constraints
related to the terrain.  Furthermore, we do not account for the
influence of noise on the wind farm's surroundings. The wind farm
literature treats the more complex farm models that capture non-flat
terrain~\cite{song2015lazy,kuo2016wind}, multi-typed
turbines~\cite{feng2017design}, noise
effects~\cite{Zhang2014,sorkhabi2016impact,yin2014multi}, and complex
objectives (including turbine installation and maintenance
costs)~\cite{lackner2007analytical,sorkhabi2018constrained}.  In this work, we 
choose to focus on a
novel approach for solving WFLO, and hence consider the most
abstract wind farm model. Extension of our framework to include such additional problem characteristics is left as future research.
%Our model can be extended, in the
%future, to include additional problem characteristics as mentioned above.

We assume that the number of identical turbines to be placed is known 
in advance and denoted by $m$. To place these $m$ turbines, 
we must consider two types of interference effects, namely \emph{proximity constraints} and \emph{wakes}.
In what follows, we explain the two effects using a toy example of placing a single turbine on an 
$8\times8$ grid (see Figure~\ref{fig:field_model}).

@@

\paragraph{Wakes}
Turbines influence each other through \emph{wake} interference effects~\cite{jensen1983note,shakoor2016wake}.
%that are
%referred to as \emph{wakes}~\cite{jensen1983note,shakoor2016wake}.
Wakes cause reduction of the effective power produced by a turbine due
to upstream turbines that change the airflow dynamics. In
Figure~\ref{fig:field_model}, the wind is assumed to be
blowing from left to right. The wake effect due to placing the turbine
expands as shown in the diagram.
%: it will influence only turbines that
%are placed in the orange cells.  In other words,
If a turbine is placed in the orange area, its effective power is
reduced due to the upstream turbine.
% that we placed in the $8\times8$ farm.
Moreover, a given turbine may be placed downstream from multiple turbines and hence
multiple wakes must be taken into account for each location.  The
reduction of power due to the wake effect is determined by three
parameters: (1) the distance between the upstream turbine and the cell
for which we want to compute the effect (denoted by $x$ in
Figure~\ref{fig:field_model}), (2) the radius of the cone's opening at distance $x$ 
(denoted by $r$ in Figure~\ref{fig:field_model}), and (3) the
combination of several wind regimes where a wind regime consists of 
the wind's free speed and direction.  The wake effect can be pre-computed
for each pair of cells since we know the distance ($x$), the turbine
specifications and terrain conditions (that jointly dictate $r$), and the
probabilistic behavior of wind regimes~\cite{Zhang2014}.

\paragraph{Effective Power Calculations in the Presence of Wakes} 
Let $\mathcal{D}$ be the set of wind regimes.
%where every regime, $d\in \mathcal{D}$, is a combination of wind direction and free speed. 
The probability of wind regime 
$d \in \mathcal{D}$ is given by $p_d$ with $\sum_{d \in \mathcal{D}}^{} p_d = 1$. In the experimental section (Section~\ref{sec:eval}),
we provide one example of a probability distribution 
function over $36$ wind regimes (Figure~\ref{fig:prob_wind}). In that example, 
a regime $d=(10^\circ, 12 \mbox{ km/h})$ corresponds to a north-easterly wind blowing at a free velocity of $12\mbox{ km/h}$. The
probability of this regime is $0.008$.

@@

\paragraph{Exact Optimization Methods} 
The WFLO problem can be solved  
to optimality, using
techniques from operations research (e.g., mixed-integer linear programming (MILP), quadratic programming (QP), and constraint programming (CP))~\cite{Zhang2014,turner2014new,donovan2005wind}.
These methods are \emph{exact}, in the sense that they guarantee that the returned solution is indeed
the best solution one can obtain for the given problem. Donovan was the first to solve WFLO using a
MILP approach based on the LS energy expression~\cite{donovan2005wind}. Turner et al.\ \cite{turner2014new} approximated the SS energy expression using a quadratic and a linear approximation. The two resulting methods performed well when
compared to existing approximate and exact techniques.
Another non-linear optimization approach \cite{ulku2019new} relaxed the binary placement variables to continuous values in $[0,1]$
and approximated the constraints such that the resulting variables yield an integer placement.
Their approach provides another approximation to the WFLO problem considered by Turner et al.\ \cite{turner2014new}.

@@

While the main advantage of exact techniques
is that they provide guarantees on solution quality, they may take a very long time to  find an optimal solution. 
In fact, Section~\ref{sec:QUBO4WFLO} shows formally that WFLO is computationally hard, a result
that we have not yet encountered in the WFLO literature. Therefore,
when one considers solving multiple wind farm design problems, if a fast solution is of essence,
one may turn to approximate optimization techniques.

\paragraph{Approximate Methods} In contrast to exact methods, approximation algorithms often provide   
quick and, hopefully, sufficiently accurate solutions. Several works have solved 
WFLO using evolutionary (genetic) approaches~\cite{MOSETTI1994105,gonzalez2010optimization,grady2005placement}. 
%These methods use two principles (cross-over and mutation)
%to change existing solutions and improve them over time.
These methods do not guarantee termination 
(i.e., they may run forever without achieving a quality criteria), and even when they do terminate (in practical cases) they do
not guarantee optimal solutions. Hence, one must often
introduce non-quality related stopping criteria, 
thus compromising on the quality of the obtained solution~\cite{davis1991handbook}.
Furthermore, it is awkward to introduce constraints into genetic methods (see e.g., \cite{sorkhabi2018constrained}), leading to potentially costly feasibility checks and problem-specific repair operations for every solution found.

The first attempt to address these limitations 
was made using a local search approach~\cite{ozturk2004heuristic}. 
The solution does not guarantee optimality, 
yet it circumvents the termination and feasibility checking issues of the evolutionary
methods. The local search approach attempts to apply either move, remove or add actions given an incumbent 
feasible solution. When a turbine is moved, removed or added, the new solution is evaluated. The procedure terminates 
when a stopping criteria is reached.
The main limitation of this approach 
%~\cite{ozturk2004heuristic}
is that it cannot perform an action 
that leads to an infeasible state and, therefore, often quickly converges to a sub-optimal solution (local optimum) 
that can potentially be much lower quality than the globally optimal solution~\cite{rivas2009solving}.

@@

In our work, we also use an annealing based approach (similar to Rivas et al.\ \cite{rivas2009solving}). However,
our methodology is generic since it is based on a declarative problem model and
on the ability of specialized hardware to solve (non-ad-hoc) quadratic problems.

@@

A benchmark study comparing state-of-the-art exact approaches \cite{Zhang2014} to 
approximate methods found that the two perform in a comparable fashion, without large differences between
the attained energy~\cite{yang2019simulated}. Exact methods tend to perform slightly better
in smaller WFLO benchmarks, while simulated annealing methods dominated for larger standard WFLO benchmarks. 
These results lead to the conclusion that both types of methods (exact and approximate) should be considered. Therefore, we introduce a unified paradigm, namely quadratic programming, that enables
both exact and approximate solutions to WFLO. Further, formulating WFLO using quadratic programming enables us to employ 
recent advanced exact software solutions (e.g., Gurobi optimizer) and approximate hardware developments 
(e.g., Fujitsu's Digital Annealer).
%Thus, 
%our approach `enjoys' the best of both worlds (the exact and the approximate).

@@

We start by presenting the inputs and decision variables into our optimization approaches.
Subsequently, we introduce the two quadratic formulations of WFLO and 
prove its computational complexity. Lastly, we discuss the details of existing
software and hardware solutions that can be used to solve the two formulations.

@@

\subsubsection{QCOP solving methods}
\label{quadratic}
\qcls{} is a strictly convex integer 
quadratically constrained optimization problem (IQCOP), which is notorious for its computational complexity~\cite{van1981another}.
%In fact, we proved in Section~\ref{sec:computational} that WFL-QCOP is $\mathcal{NP}$-hard.
The common exact methods for solving
IQCOPs are tree search \cite{land2010automatic} algorithms with a variety of sophisticated techniques to limit the search necessary to find and prove optimality.  In operations research, the generic exact approach for solving IQCOPs is the branch-and-cut based mixed integer programming (MIP) \cite{bonami2008algorithmic}. A MIP solver, such as the one we use (Gurobi), uses a linear or quadratic continuous relaxation to efficiently compute a bound on the objective function. This relaxation is used to generate cuts, implied constraints that tighten the relaxation, to identify sub-trees that can be discarded without search because they are guaranteed not to contain a solution better than the current incumbent, and to drive the heuristic branching that systematically searches for a solution. Typically, this process of cut generation, reasoning about sub-trees, and branching is performed at each node in the search tree.

Commercial MIP solvers have seen extensive development over the past 20 years and hardware-independent algorithmic improvements have delivered orders of magnitude improved performance \cite{Bixby07a}. More recently the development of such solvers has focused on representing and solving quadratic problems such as \qcls{}~\cite{Furini19a}.
%By solving a relaxed problem to bound the objective value at each node and strengthening the relaxed problem with cutting planes to restrict the feasible region further, MIP solvers has been proved effective for attacking IQCOPs \cite{junger200950}.

>>>>

\begin{abstract} 
%Wind farms play an increasingly important
%role as a source of renewable energy and 
  A key aspect of the design of a wind farm is the wind farm layout optimization (WFLO) problem: given a wind farm site and information about the wind patterns, the problem is to decide the location of individual wind turbines to maximize energy production subject to proximity restrictions and wake-based interference between turbines. Given the pairwise wake interactions, it is natural
%While a number of heuristic and exact techniques have been proposed to solve WFLO, it is common
to model the energy objective as a quadratic function as, indeed, has been done in some existing optimization approaches. However, state-of-the-art solutions often trade-off between speed in producing designs and quality in terms of finding and proving optimal solutions. In this work, we aim to find a balanced approach to obtain WFLO solutions quickly for interactive design \emph{and} solve the problem to optimality when quality is more important. To this end, we exploit recent advances in optimization hardware and software that target quadratic constraints: commercial mixed integer linear solvers have been extended to address some quadratic problems and nascent specialized hardware, including quantum computing systems, have focused on solving quadratic unconstrained binary optimization (QUBO) problems. 
%It is a key problem 
%when designing a new wind farm, as it has a direct influence
%on the amount of the total produced energy. 
%In recent years there have been advances in both software technologies (e.g., the extension of commercial mixed integer linear programming solvers to handle quadratic problems) and nascent hardware development (e.g., quantum computing chips and specialized classical chips) that have relevance for the WFLO problem. In both cases, the advances have 
%Recent years saw 
%advances in software and hardware optimization technologies, such as Gurobi's quadratic solver 
%and Fujitsu's digital annealer (DA), respectively. 
%These technologies create an opportunity to improve 
%over current wind farm optimization solutions
%both in terms of solution quality and solve time. 
%In order to exploit these technologies, 
%we must represent the WFLO problem as 
%a quadratic optimization model. Specifically, in order 
%to apply a software-based optimization
%solution (Gurobi), 
We 
introduce two novel quadratic programming models for WFLO: a quadratic constrained optimization problem (QCOP) with binary decision variables and a QUBO.
%Subsequently, we provide the unconstrained counterpart of the QCOP, 
%namely the quadratic unconstrained binary optimization (QUBO), which  
%is the underlying representation for quantum-inspired 
%optimization hardware such as the DA. 
A thorough numerical 
evaluation using a commercial solver and specialized QUBO hardware
%demonstrates the strengths of the two models. 
%Our results
show that our quadratic framework achieves fast, high-quality solutions 
that
improve the state of the art and strike a balance between speed and quality. In particular, the QUBO model delivers high quality solutions in a few seconds while the QCOP model can be used to find better solutions and provide quality guarantees over a longer run-time.
\end{abstract}
\begin{keyword}
	Wind Farm Layout Optimization (WFLO) \sep Quadratic Programming (QP) \sep Quadratic Unconstrained Binary Optimization (QUBO) \sep Digital Annealer
\end{keyword}

@@

\section{Introduction}

The goal in the wind farm layout optimization (WFLO)
problem~\cite{MOSETTI1994105} is to locate wind turbines within a
predefined area to maximize the total power harvested from the wind
stream.  While physical considerations based on the diameter of the
blades prevent turbines being placed too close to each other, turbine
location decisions influence the total power production due to
interference effects (i.e., wakes) generated by upstream turbines.
These inter-turbine effects are best captured by Jensen's wake model~\cite{jensen1983note} when superimposed
using a sum-of-squares expression~\cite{Zhang2014}. However,
this wake modeling yields intractable optimization problems
that must be approximated using either quadratic or linear functions~\cite{turner2014new}.
An alternative formulation of WFLO relies on a less accurate wake model, using linear superpositioning
to capture multiple wakes~\cite{donovan2005wind}.
The latter can be captured using quadratic
constraints and objective functions. However, exact optimization
approaches that have been proposed for the alternative WFLO formulation~\cite{Zhang2014,kuo2016wind,donovan2005wind,fagerfjall2010optimizing,archer2011wind,sorkhabi2018constrained}
represent the quadratic relations using approximate linear models,
%that approximate wake dependencies,
as solving quadratic programming problems
is computationally challenging compared to linear models.

@@

In this work, we introduce two novel quadratic WFLO models with the goal 
of combining efficiency, physical farm representation, and optimality. The first
model represents WFLO as a quadratic constrained optimization
problem (QCOP).  The QCOP can be solved using state-of-the-art mathematical software
that has been recently extended for such problems, e.g., Gurobi~\cite{gurobi}.  Our
second model is a quadratic unconstrained binary optimization
(QUBO) representation of the WFLO problem.  Such a formulation enables the use of
nascent specialized optimization hardware tailored to solve QUBOs: we use Fujitsu's Digital Annealer (DA) \cite{aramon2019physics}. By employing the two types of advances (software and hardware) for solving
quadratic problems, 
we are able to quickly find high quality solutions in terms of the total energy. Moreover, our approach is based
on a physical model of the wind farm, taking into account
exact calculations of the total energy and the wake effects. 
These two models together achieve a balance between efficiency and effectiveness of the resulting design, in a model-based fashion.
%, a hardware-based
%QUBO solver, to demonstrate the strength of our approach.

@@

Our experiments are focused on the performance of the two approaches
compared to 
existing optimization methods. To this end,
we experiment with 12 standard WFLO benchmark problems (c.f.~\cite{turner2014new})
and show that solving quadratic WFLO models
using Gurobi and the DA
achieves state-of-the-art performance.
%, often outperforming existing solutions when provided sufficient run-time.

%Furthermore, we observe that the DA finds solutions of
%similar quality to software-based approaches but with two orders of
%magnitude shorter run-time.

%The ability to quickly obtain competitive
%solutions on large WFLO instances can be employed to solve multiple
%wind farm design problems.

@@

We start this section by presenting the physical model of a wind farm
%used throughout the paper.
%Our wind farm model is
based on the common
notation and description of Zhang et al.\ \cite{Zhang2014}.  Subsequently,
we review the literature on existing mathematical optimization
methods for solving WFLO, focusing on similarities and
differences with our approach.

@@

%\paragraph{Wind Farm as a Grid}
We consider a square wind farm that we model as an $n\times n$
grid with $n$ being the number of cells on each axis.  The cells have
an area of $c^2$ (in square meters), i.e., if $c=100$ meters and
$n=20$, the farm area will be $2 \mbox{ km} \times 2 \mbox{ km}$.  We
assume that the terrain is flat and that we have no constraints
related to the terrain.  Furthermore, we do not account for the
influence of noise on the wind farm's surroundings. The wind farm
literature treats the more complex farm models that capture non-flat
terrain~\cite{song2015lazy,kuo2016wind}, multi-typed
turbines~\cite{feng2017design}, noise
effects~\cite{Zhang2014,sorkhabi2016impact,yin2014multi}, and complex
objectives (including turbine installation and maintenance
costs)~\cite{lackner2007analytical,sorkhabi2018constrained}.  In this work, we 
choose to focus on a
novel approach for solving WFLO, and hence consider a more
abstract wind farm model. Extension of our framework to include such additional problem characteristics is left as future research.
%Our model can be extended, in the
%future, to include additional problem characteristics as mentioned above.

We assume that the number of identical turbines to be placed is known 
in advance and denoted by $m$. To place these $m$ turbines, 
we must consider two types of interference effect: \emph{proximity constraints} and \emph{wakes}.
In what follows, we explain the two effects using a toy example with an 
$8\times8$ grid (see Figure~\ref{fig:field_model}).

@@

\paragraph{Wakes}
Turbines influence each other through \emph{wake} interference effects that reduce the effective power produced by a turbine due to upstream turbines that change the airflow dynamics~\cite{jensen1983note,shakoor2016wake}.
%that are
%referred to as \emph{wakes}~\cite{jensen1983note,shakoor2016wake}.
 In
Figure~\ref{fig:field_model}, the wind is assumed to be
blowing from left to right. The wake effect due to placing the turbine
expands as shown in the diagram.
%: it will influence only turbines that
%are placed in the orange cells.  In other words,
If a turbine is placed in the orange area, its effective power is
reduced due to the upstream turbine.
% that we placed in the $8\times8$ farm.
Moreover, a given turbine may be placed downstream from multiple turbines and hence
multiple wakes must be taken into account for each location.  The
reduction of power due to the wake effect is determined by three
parameters: (1) the distance between the upstream turbine and the cell
for which we want to compute the effect (denoted by $x$ in
Figure~\ref{fig:field_model}), (2) the radius of the cone's opening at distance $x$ 
(denoted by $r$ in Figure~\ref{fig:field_model}), and (3) the
combination of several wind regimes where a wind regime consists of 
the wind's free speed and direction.  The wake effect can be pre-computed
for each pair of cells since we know the distance ($x$), the turbine
specifications and terrain conditions (that jointly dictate $r$), and the
probabilistic behavior of wind regimes~\cite{Zhang2014}.

\paragraph{Effective Power Calculations in the Presence of Wakes} 
Let $\mathcal{D}$ be the set of wind regimes.
%where every regime, $d\in \mathcal{D}$, is a combination of wind direction and free speed. 
The probability of wind regime 
$d \in \mathcal{D}$ is given by $p_d$ with $\sum_{d \in \mathcal{D}}^{} p_d = 1$. In the experimental section (Section~\ref{sec:eval}),
we provide one example of a probability distribution 
function over $36$ wind regimes (Figure~\ref{fig:prob_wind}). In that example, 
one regime $d=(10^\circ, 12 \mbox{ km/h})$ corresponds to a north-easterly wind blowing at a free velocity of $12\mbox{ km/h}$. The
probability of this regime is $0.008$.

@@

\paragraph{Exact Optimization Methods} 
The WFLO problem can be solved  
to optimality, using
techniques from operations research (e.g., mixed-integer linear programming (MILP), quadratic programming (QP), and constraint programming (CP))~\cite{Zhang2014,turner2014new,donovan2005wind}.
These methods are \emph{exact}, in the sense that they guarantee that the returned solution is indeed
the best solution one can obtain for the given model of the problem. Donovan was the first to solve WFLO using a
MILP approach based on the LS energy expression~\cite{donovan2005wind}. Turner et al.\ \cite{turner2014new} approximated the SS energy expression using a quadratic and a linear approximation. The two resulting methods performed well when
compared to existing approximate and exact techniques.
Another non-linear optimization approach \cite{ulku2019new} relaxed the binary placement variables to continuous values in $[0,1]$
and approximated the constraints such that the resulting variables yield an integer placement.
Their approach provides another approximation to the WFLO problem considered by Turner et al.\ \cite{turner2014new}.

@@

While the main advantage of exact techniques
is that they provide guarantees on solution quality, they may take a very long time to  find an optimal solution. 
In fact, Section~\ref{sec:QUBO4WFLO} shows formally that WFLO is computationally hard, a result
that is not surprising but that has apparently not explicitly appeared in the WFLO literature. Therefore,
when one considers solving multiple wind farm design problems, if a fast solution is of essence,
one may turn to approximate optimization techniques.

\paragraph{Approximate Methods} In contrast to exact methods, approximation algorithms often provide   
quick and, hopefully, sufficiently high-quality solutions. Several works have solved 
WFLO using evolutionary (genetic) approaches~\cite{MOSETTI1994105,gonzalez2010optimization,grady2005placement}. 
%These methods use two principles (cross-over and mutation)
%to change existing solutions and improve them over time.
These methods do not guarantee termination 
(i.e., they may run forever without achieving a quality criteria), and even when they do terminate they do
not guarantee optimal solutions. Hence, one must often
introduce non-quality related stopping criteria, 
compromising the quality of the obtained solution~\cite{davis1991handbook}.
Furthermore, it is awkward to introduce constraints into genetic methods (see e.g., \cite{sorkhabi2018constrained}), leading to potentially costly feasibility checks and problem-specific repair operations for every solution found.

The first attempt to address these limitations 
was made using a local search approach~\cite{ozturk2004heuristic}. 
The approach does not guarantee optimality, 
yet it circumvents the termination and feasibility checking issues of the evolutionary
methods. The local search attempts to apply either move, remove or add actions given an incumbent 
feasible solution. When a turbine is moved, removed or added, the new solution is evaluated. The procedure terminates 
when a stopping criteria is reached.
The main limitation of this approach 
%~\cite{ozturk2004heuristic}
is that it cannot perform an action 
that leads to an infeasible state and, therefore, often quickly converges to a sub-optimal solution (local optimum) 
that can potentially be much lower quality than the globally optimal solution~\cite{rivas2009solving}.

@@

In our work, we also use an annealing based approach (similar to Rivas et al.\ \cite{rivas2009solving}). However,
our methodology is generic since it is based on a declarative problem model and
on the ability of specialized hardware to solve %(non-ad-hoc)
quadratic problems.

@@

A benchmark study comparing state-of-the-art exact approaches \cite{Zhang2014} to 
approximate methods found that the two perform in a comparable fashion, without large differences between
the attained energy~\cite{yang2019simulated}. Exact methods tend to perform slightly better
in smaller WFLO benchmarks, while simulated annealing methods dominated for larger standard WFLO benchmarks. 
These results led to the conclusion that both types of methods (exact and approximate) should be considered. Therefore, we introduce a unified paradigm, namely quadratic programming, that enables
both exact and approximate solutions to WFLO. Further, formulating WFLO using quadratic programming enables us to employ 
recent advanced exact software solutions (e.g., Gurobi optimizer) and approximate hardware developments 
(e.g., Fujitsu's Digital Annealer).
%Thus, 
%our approach `enjoys' the best of both worlds (the exact and the approximate).

@@

We start by presenting the inputs and decision variables of our optimization approaches.
Subsequently, we introduce the two quadratic formulations of WFLO and 
prove its computational complexity. Lastly, we discuss the details of existing
software and hardware solutions that can be used to solve the two formulations.

@@

\subsubsection{QCOP solving methods}
\label{quadratic}
\qcls{} is a strictly convex integer 
quadratically constrained optimization problem (IQCOP), which is notorious for its computational complexity~\cite{van1981another}.
%In fact, we proved in Section~\ref{sec:computational} that WFL-QCOP is $\mathcal{NP}$-hard.
The common exact methods for solving
IQCOPs are tree search \cite{land2010automatic} algorithms with a variety of sophisticated techniques to limit the search necessary to find and prove optimality.  In operations research, the generic exact approach for solving IQCOPs is the branch-and-cut based mixed integer programming (MIP) \cite{bonami2008algorithmic}.

%A MIP solver, such as the one we use (Gurobi), uses a linear or quadratic continuous relaxation to efficiently compute a bound on the objective function. This relaxation is used to generate cuts, implied constraints that tighten the relaxation, to identify sub-trees that can be discarded without search because they are guaranteed not to contain a solution better than the current incumbent, and to drive the heuristic branching that systematically searches for a solution. %Typically, this process of cut generation, reasoning about sub-trees, and branching is performed at each node in the search tree.

Commercial MIP solvers have seen extensive development over the past 20 years with hardware-independent algorithmic improvements have delivered orders of magnitude improved performance \cite{Bixby07a}. More recently the development of such solvers has focused on representing and solving quadratic problems such as \qcls{}~\cite{Furini19a}.
%By solving a relaxed problem to bound the objective value at each node and strengthening the relaxed problem with cutting planes to restrict the feasible region further, MIP solvers has been proved effective for attacking IQCOPs \cite{junger200950}.
